Unsupervised training of maximum-entropy models for lexical selection in rule-based machine translation

### 2015-04-13 2015-eamt-melsa/2015-eamt-melsa.tex
some fixes
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59731 72bbbca6-d526-0410-a7d9-f06f51895060

\subsection{Prior work}
%@@@ \comment{This section should be ironed out a bit.}  
%In their pioneering work, 
\cite{dagan1994word} used the term \emph{word sense disambiguation} to
refer to what is actually target word selection\todofsm{Should we call
  it lexical selection instead of target word selection?} in MT; they
used a parser to identify syntactic relations such as subject--object
or subject--verb.  After generating all the possible translations for
a given input sentence using an ambiguous bilingual dictionary, they
extract the syntactic tuples from the TL and count the frequency in a
previously-trained TL model of tuples.  They use maximum-likelihood
estimation to calculate the probability that a given TL tuple is the
translation of a given SL tuple, with an automatically determined
confidence threshold.

@@

Let the probability of a word $t$ being the translation of a word $s$
in a SL context $c$ be $p_s(t|c)$. In principle, this value could be
estimated directly from the available corpora for every combination of
$(s, t, c)$.  This would however present two questions: (1) how should
the relevant contexts be chosen? and (2) what should be done when $(s,
t, c)$ is not found in the corpus?  A maximum-entropy model answers
both of these questions. It allows the contexts that we consider to be
linguistically interesting to be defined \emph{a priori} and then
integrate these seamlessly into a probabilistic model
\citep{Manning99b}.  In answer to the second question, a
maximum-entropy model maximise the entropy subject to match the
expected counts of the designed features with those found in the
training data. \todofsm{Solved R3: ``In answer to the second question, a
  maximum-entropy model assumes nothing about what is not present in
  the training data. That is, if there is no information in the
  training data, then it assumes that all outcomes —that is, all
  possible translations— are equally likely'' can be applied to a
  uniform distribution for a discrete variable. I think the authors
  are trying to restate the main property of the maximum entropy
  model, i.e.  maximum entropy model maximize the entropy (which is
  the uniformness of the distribution) subject to match the expected
  counts of the designed features with those found in the training
  data.  }
%assumes nothing about what is not present in the
%training data. 
That is, if there is no information in the training data, then it
assumes that all outcomes ---that is, all possible translations--- are
equally likely.  As previously mentioned, the principle of maximum
entropy has been applied to the problem of lexical selection before;
in particular, \cite{berger1996} cast the problem of lexical selection
in statistical MT as a classification problem. They learn a separate
maximum-entropy classifier for each SL word form,
%\comment{[:we do too, integrate this later]} 
using SL context to distinguish between possible translations. These
classifiers are then incorporated into the translation model of their
word-based statistical MT system.
%\comment{[how is our work different?]}  
In their approach, a classifier consists of a set of binary feature
functions and corresponding weights for each feature.  In both
\cite{berger1996} and our method, features are defined in the form
$h_k^{s}(t, c)$, where $t$ is a translation, and $c$ is a SL
context. One difference is that \cite{berger1996} take $s$, $t$ and
$c$ to be based on word forms, whereas in our method they are based on
lemma forms.  An example would be the following feature where the
Spanish word \emph{pez} (`fish' as a living animal) is seen as the
translation of \emph{arrain} (`fish') in the context \emph{arrain
  handi} `big fish' and would therefore be defined as:
%\comment{FSM: I have introduced the index $k$
%  because several features are defined for a SL word $s$.} 
\begin{equation}
h_\mathrm{+handi}^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
                 ~~1 & \mbox{if } 
\left\{ \begin{array}{c}
t~ =~ \textit{pez}~ \\ \mathrm{and}~ \\ \textit{handi}~ \mathrm{follows}~ \textit{arrain} \end{array} \right. \\
                 ~~0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:feat-func}
\end{equation}
This feature considers a context of zero words to the left of the
problem word and one word (+ \emph{handi}) to the right of it.\todomlf{Removed \(k\) and used a more informative subscript}

@@

As a result of training, each of the \(n_F\) features \(h^s_k(t,c)\) in the classifier is assigned a
weight $\lambda_k^s$. \todomlf{Reworded to define \(n_F\)} Combining these weights of active features
as in equation~(\ref{eq:max-ent-sum}) yields the probability of a
translation $t$ for word $s$ in context $c$.

\begin{equation}
p_s(t|c) = \frac{1}{Z^s(c)} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
\label{eq:max-ent-sum}
\end{equation}
\todomlf{R3 complains: `` Probability notation is confusing. I would
  suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will
  simplify notation. For instance instead of \(Z^s (c)\) we would have
  \(Z(c,s)\).''. Ignore or take onboard? FSM: Mention that we follow
  the notation of Berger?}

>>>>

\subsection{Prior work}
%@@@ \comment{This section should be ironed out a bit.}  
%In their pioneering work, 
\cite{dagan1994word} used the term \emph{word sense disambiguation} to
refer to what is actually lexical selection in MT; they
used a parser to identify syntactic relations such as subject--object
or subject--verb.  After generating all the possible translations for
a given input sentence using an ambiguous bilingual dictionary, they
extract the syntactic tuples from the TL and count the frequency in a
previously-trained TL model of tuples.  They use maximum-likelihood
estimation to calculate the probability that a given TL tuple is the
translation of a given SL tuple, with an automatically determined
confidence threshold.

@@

Let the probability of a word $t$ being the translation of a word $s$
in a SL context $c$ be $p_s(t|c)$. In principle, this value could be
estimated directly from the available corpora for every combination of
$(s, t, c)$.  This would however present two questions: (1) how should
the relevant contexts be chosen? and (2) what should be done when $(s,
t, c)$ is not found in the corpus?  A maximum-entropy model answers
both of these questions. It allows the contexts that we consider to be
linguistically interesting to be defined \emph{a priori} and then
integrate these seamlessly into a probabilistic model
\citep{Manning99b}.  In answer to the second question, a
maximum-entropy model maximise the entropy subject to match the
expected counts of the designed features with those found in the
training data. \todofsm{Solved R3: ``In answer to the second question, a
  maximum-entropy model assumes nothing about what is not present in
  the training data. That is, if there is no information in the
  training data, then it assumes that all outcomes —that is, all
  possible translations— are equally likely'' can be applied to a
  uniform distribution for a discrete variable. I think the authors
  are trying to restate the main property of the maximum entropy
  model, i.e.  maximum entropy model maximize the entropy (which is
  the uniformness of the distribution) subject to match the expected
  counts of the designed features with those found in the training
  data.  }
%assumes nothing about what is not present in the
%training data. 
That is, if there is no information in the training data, then it
assumes that all outcomes ---that is, all possible translations--- are
equally likely.  As previously mentioned, the principle of maximum
entropy has been applied to the problem of lexical selection before;
in particular, \cite{berger1996} cast the problem of lexical selection
in statistical MT as a classification problem. They learn a separate
maximum-entropy classifier for each SL word form,
%\comment{[:we do too, integrate this later]} 
using SL context to distinguish between possible translations. These
classifiers are then incorporated into the translation model of their
word-based statistical MT system.
%\comment{[how is our work different?]}  
In their approach, a classifier consists of a set of binary feature
functions and corresponding weights for each feature.  In both
\cite{berger1996} and our method, features are defined in the form
$h_k^{s}(t, c)$,\footnote{We follow the notation of \cite{berger1996}} where $t$ is a translation, and $c$ is a SL
context. One difference is that \cite{berger1996} take $s$, $t$ and
$c$ to be based on word forms, whereas in our method they are based on
lemma forms.  An example would be the following feature where the
Spanish word \emph{pez} (`fish' as a living animal) is seen as the
translation of \emph{arrain} (`fish') in the context \emph{arrain
  handi} `big fish' and would therefore be defined as:
%\comment{FSM: I have introduced the index $k$
%  because several features are defined for a SL word $s$.} 
\begin{equation}
h_\mathrm{+handi}^{\mathrm{arrain}}(t, c) = \left \{ \begin{matrix}
                 ~~1 & \mbox{if } 
\left\{ \begin{array}{c}
t~ =~ \textit{pez}~ \\ \mathrm{and}~ \\ \textit{handi}~ \mathrm{follows}~ \textit{arrain} \end{array} \right. \\
                 ~~0 & \mbox{otherwise} \end{matrix} \right.
\label{eq:feat-func}
\end{equation}
This feature considers a context of zero words to the left of the
problem word and one word (+ \emph{handi}) to the right of it.

@@

As a result of training, each of the \(n_F\) features \(h^s_k(t,c)\) in the classifier is assigned a
weight $\lambda_k^s$. Combining these weights of active features
as in equation~(\ref{eq:max-ent-sum}) yields the probability of a
translation $t$ for word $s$ in context $c$.

\begin{equation}
p_s(t|c) = \frac{1}{Z^s(c)} \exp{ \sum_{k=1}^{n_F} \lambda_k^s h_k^s(t, c)}
\label{eq:max-ent-sum}
\end{equation}
%\todomlf{R3 complains: `` Probability notation is confusing. I would
%  suggest the use of \(p(t | c,s)\) instead of \(p_s(t|c)\), that will
%  simplify notation. For instance instead of \(Z^s (c)\) we would have
%  \(Z(c,s)\).''. Ignore or take onboard? FSM: Mention that we follow
%  the notation of Berger?}
